My title

### 2023-03-21 outputs/paper/paper.qmd
added bleachers data, edited the data sections

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: html
number-sections: true
bibliography: references.bib
---

@@

@@

# Appendix {-}

>>>>

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Michaela Drouillard
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: html
number-sections: true
bibliography: references.bib
---

@@

I pulled the data using the Spotify API, which I accessed using the `spotifyr` package in R. I used the `get_artist_audio_features` and `get_track_audio_analysis functions` function to acquire data on Lorde, Taylor Swift, St. Vincent, Lana Del Rey, The Chicks, Florence and the Machine, and Bleachers discographies.

### Artist Selection

I chose these artists because they've produced multiple albums, and at least one album with Antonoff, as opposed to just individual songs. I chose the Chicks because I figured that collaborating with Antonoff after an extended gap in a longer spanning career than the other artists might offer some interesting results. The data points at beginning of their career reflect music that existed in a different time from the other artists in this dataset.

### Data Cleaning

I joined these data sets using the `rbind` function in base R, and created an additional variable called "jack", which contains a 1 if Jack Antonoff produced or co-produced the song, and a 0 if he had nothing to do with it. I got the information on which songs Antonoff did or did not produce from the "Jack Antonoff Production Discography" Wikipedia page.

### Spotify's API

So what do any one these features mean?

The Spotify API returns JSON metadata directly from the Spotify Data Catalogue.

#### Audio Features Dataset

The `get_artist_audio_features` provides a general profile of each song, according to variables developped by spotify, which makes for simple comparisons across artists.

However, it's hard to gauge exactly what variables like "danceability" or "valence" mean, even using the documentation. For instance, "valence" is described as "A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track. Tracks with high valence sound more positive (e.g. happy, cheerful, euphoric), while tracks with low valence sound more negative (e.g. sad, depressed, angry)", and danceability as "how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity. A value of 0.0 is least danceable and 1.0 is most danceable". Neither of these definitions fully cover what metrics and thresholds go into something like "danceability". However, whatever these features capture is easier to compare across artists in visualizations than more granular details about bars, beats, duration, and tatums -- a word I just learned, which refers to "the lowest regular pulse train that a listener intuitively infers from the timing of perceived musical events". The full documentation for each variable included in this study is available in the Appendix.

The final audio features in this study data set contains the variables: "artist_name", "track_name", "energy", "danceability", "energy", "key", "loudness", "mode", "speechiness", "acousticness", "instrumentalness", "liveness", "valence", "tempo", "jack".

```{r}
library(kableExtra)
df |>
  knitr::kable()

#### Audio Analysis Dataset

The `get_track_audio_analysis` offers a more granular look at each track. The variables included are arrays of objects, which are the kinds of things that are a bit harder to wrap your mind around if you were to visualize each one across each artist. They include: track, bars, beats, sections, segments, tatums, and meta, which includes information on which platforms and Analyzer versions were used to analyze the track, along with metadata on the status and duration of the analysis.

This data will be used when building a logistic regression model to determine the character and predictability of Jack Antonoff's music.

@@

# Appendix {.unnumbered}

### `get_artist_audio_features` documentation:

-   **acousticness**: A confidence measure from 0.0 to 1.0 of whether the track is acoustic. 1.0 represents high confidence the track is acoustic.

-   **danceability**: Danceability describes how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity. A value of 0.0 is least danceable and 1.0 is most danceable.

-   **energy**: Energy is a measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity. Typically, energetic tracks feel fast, loud, and noisy. For example, death metal has high energy, while a Bach prelude scores low on the scale. Perceptual features contributing to this attribute include dynamic range, perceived loudness, timbre, onset rate, and general entropy.

-   **instrumentalness**: Predicts whether a track contains no vocals. "Ooh" and "aah" sounds are treated as instrumental in this context. Rap or spoken word tracks are clearly "vocal". The closer the instrumentalness value is to 1.0, the greater likelihood the track contains no vocal content. Values above 0.5 are intended to represent instrumental tracks, but confidence is higher as the value approaches 1.0.

-   **key**: The key the track is in. Integers map to pitches using standard [Pitch Class notation](https://en.wikipedia.org/wiki/Pitch_class). E.g. 0 = C, 1 = C♯/D♭, 2 = D, and so on. If no key was detected, the value is -1.

-   **liveness**: Detects the presence of an audience in the recording. Higher liveness values represent an increased probability that the track was performed live. A value above 0.8 provides strong likelihood that the track is live.

-   **loudness:** The overall loudness of a track in decibels (dB). Loudness values are averaged across the entire track and are useful for comparing relative loudness of tracks. Loudness is the quality of a sound that is the primary psychological correlate of physical strength (amplitude). Values typically range between -60 and 0 db.

-   **mode:** Mode indicates the modality (major or minor) of a track, the type of scale from which its melodic content is derived. Major is represented by 1 and minor is 0.

-   **tempo:** The overall estimated tempo of a track in beats per minute (BPM). In musical terminology, tempo is the speed or pace of a given piece and derives directly from the average beat duration.

-   **valence:** A measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track. Tracks with high valence sound more positive (e.g. happy, cheerful, euphoric), while tracks with low valence sound more negative (e.g. sad, depressed, angry).
